Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-06-10 pirouette_supplement.tex
Reviewed interprations in appendix

\new{
  The resulting error distributions are shown in Fig. \ref{fig:replicate_trees}. 
  We present results for cases where (1) the generative model has been used 
  or (2) the model with highest evidence has been selected for the inference.
  From the plots we can see that in both cases there is a clear distinction 
  between the two distributions (true and twin), suggesting that the inference 
  models that have been used cannot capture in an accurate way the features of 
  the diversity-dependent tree prior used to simulate the original trees. 
}

@@

\new{
We show in figures \ref{fig:example_6_taxa}, \ref{fig:example_12_taxa}, \ref{fig:example_24_taxa}, \ref{fig:example_32_taxa} and \ref{fig:example_40_taxa} what are the errors obtained when starting from phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa. Again we can see that in each case errors tend to be greater in the true distribution than in the twin distribution, confirming the result of the previous paragraph. However we cannot detect a clear trend between the number of taxa of the original phylogeny and the extent of such discrepancy. Due to long computational times we have been able to provide data only for a limited amount of replicates. We cannot exclude that a clearer trend could be observed increasing the number of replicates.
}

@@

\new{
From figures \ref{fig:example_500_nucleotides}, \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} we can observe a that the discrepancy between the true and twin error distributions tend to become smaller as the number of nucleotides increase. This occurred for both the generative and best candidate cases. 
}
\giovanni{richel do you have an idea why this happens? Also, do we have the raw data stored somewhere? From what I can judge from the plots the median discrepancy is 0.012 for 500 nucs, 0.008 for 1000 nucs and 0.003 for 2000 nucs.}

@@

This example shows a parameterization at the correct level for the
simplest case possible.
\new{
As expected the twin and true distributions in Fig. \ref{fig:example_yule} are extremely similar for both the generative and the best candidate case.
}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
\end{figure}

\giovanni{This is the example one of the reviewer was referring to. The figure that was present before (still present in the overleaf files) looked actually quite different from the one on the repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}). I changed it with the one found there. In case you think this is wrong you can revert it back.}

\new{
Since the two models are very similar to each other (the BD model can be turned into a Yule model just by setting the extinction parameter to zero \giovanni{Cite Louca and Pennell here?}) the median discrepancy it's almost negligible. However, respect to the previous case when a Yule tree prior has been considered, here the distributions exhibit a greater difference in shapes.
}

@@

The main example uses a twin alignment that has the same number
of \new{substitution}s (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
the twin alignment uses the same mutation rate, yet is not guaranteed
to have the same number of \new{substitution}s.
\iffalse
Any difference in the error distributions can therefore also stem from a difference in the amount of information contained in the alignments.
\fi

@@

\giovanni{I am not able to give an interpretation to these figures. I would expect a trend between the difference of true and twin medians and mutation rate, but I cannot see anything. The fact that sometimes true is smaller and some other times twin is, makes me think that these results are probably not very reliable, probably due to a not sufficient number of replicates.}

>>>>

\new{
  The resulting error distributions are shown in Fig. \ref{fig:replicate_trees}. 
  We present results for cases where (1) the generative model has been used 
  or (2) the model with highest evidence has been selected for the inference.
  From the plots we can see that in both cases the two distributions 
  (true and twin) are mostly overlapping, but not everywhere.
  This suggests that the inference 
  models that have been used can to a reasonable extent capture in an accurate way 
  the features of the diversity-dependent tree prior used to simulate 
  the original trees. 
}

@@

\new{
  We show in figures \ref{fig:example_6_taxa}, \ref{fig:example_12_taxa}, 
  \ref{fig:example_24_taxa}, \ref{fig:example_32_taxa} and 
  \ref{fig:example_40_taxa} what are the errors obtained when starting from 
  phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa. Again we can see 
  that in each case errors tend to be greater in the true distribution than 
  in the twin distribution, similar to the result of the previous 
  subsection, subsection\ref{subsec:distribution}.
  In addition to this, we cannot detect a clear trend between the number of 
  taxa of the original phylogeny and the extent of such discrepancy. 
  Due to long computational times we have been able to provide data only 
  for a limited amount of replicates. 
  We cannot exclude that a clearer trend could be observed increasing the 
  number of replicates.
}

@@

\new{
  From figures \ref{fig:example_500_nucleotides}, 
  \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
  we can observe a that the discrepancy between the true and twin error 
  distributions tend to become smaller as the number of nucleotides increase. 
  This occurred for both the generative and best candidate cases.
  This follows the expectation that a prior becomes less important when
  more information becomes available.
}
\giovanni{
  Also, do we have the raw data 
  stored somewhere? From what I can judge from the plots the median discrepancy 
   is 0.012 for 500 nucs, 0.008 for 1000 nucs and 0.003 for 2000 nucs.
}
\richel{
  Sure, raw data is at
  \url{https://www.richelbilderbeek.nl/pirouette_example_19.zip},
  \url{https://www.richelbilderbeek.nl/pirouette_example_28.zip}
  and 
  \url{https://www.richelbilderbeek.nl/pirouette_example_34.zip}.
  These URLS can be found on the GitHub site.
  What would you like to with the raw data?
}

@@

This example shows a parameterization at the correct level for the
simplest case possible.

\new{
  As expected the twin and true distributions in Fig. \ref{fig:example_yule} 
  are extremely similar for both the generative and the best candidate case.
}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
  \caption{Aggregate error distributions for 10 replicates. 
  Here each true tree is generated by a BD process. 
  For the inference we used instead a Yule tree prior. This took 
  \richel{unknown} hours to compute.
}
\end{figure}

\giovanni{
  This is the example one of the reviewer was referring to. 
  The figure that was present before (still present in the overleaf files) 
  looked actually quite different from the one on the 
  repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}). 
  I changed it with the one found there. In case you think this is wrong you can revert it back.
}
\richel{ 
  Well spotted! Thanks to you, the figure is correct now
}
\new{
  Since the two models are very similar to each other (the BD model can be 
  turned into a Yule model just by setting the extinction parameter to 
  zero [\cite{nee1994reconstructed}]) the median discrepancy is
  almost negligible. 
  However, with respect to the previous 
  case (subsection \ref{subsec:simplest_correct_parameterization}), 
  where a Yule tree prior has been used, 
  the distributions here exhibit a greater difference.
}

@@

The main example uses a twin alignment that has the same number
of \new{substitution}s (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
the twin alignment uses the same mutation rate, yet is not guaranteed
to have the same number of \new{substitution}s.

@@

\giovanni{
  I am not able to give an interpretation to these figures. 
  I would expect a trend between the difference of true and twin medians 
  and mutation rate, but I cannot see anything. The fact that sometimes true 
  is smaller and some other times twin is, makes me think that these results 
  are probably not very reliable, probably due to a not sufficient number 
  of replicates.
}
\richel{
  I interpreted our curvy friends :-)
}
